---
title: "Homework 05"
author: "Ayanna Chatman"
self-contained: true
date: "2023-09-27"
output: html_document
---

## Question 1

```{r}
library(tidyverse)
exercise_data <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/visualize_data.csv")
```

I predict that people who record more exercise will have less BMI.

```{r}
cor(exercise_data$Exercise, exercise_data$BMI)
```

```{r}
ggplot(data = exercise_data,
       mapping = aes(x = Exercise, y = BMI)) +
  geom_point()
```

## Question 2

```{r}
library(causact)
 ?corruptDF
```

CPI2017 is the level of public sector corruption seen in a country/territoty in 2017 on a scale of 0-100.

HDI2017 measure of human development across different countries for 2017.

## Question 3

```{r}
ggplot(corruptDF,
       mapping = aes(x = CPI2017, y = HDI2017)) +
  geom_point()
```

Based on the scatterplot, as the corruption perceptions index score increases, so does the human development index score increases as well.

## Question 4

```{r}
ggplot(corruptDF,
       mapping = aes(x = CPI2017, y = HDI2017)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r}
ggplot(corruptDF,
       mapping = aes(x = CPI2017, y = HDI2017)) +
  geom_point() +
  geom_smooth(method = "gam")
```

When using the "lm" the line that shown is straighter, showing a more linear trend. Whereas the line that's produced when using "gam" follows the natural curve of the data, which is why I prefer the "gam" line.

## Question 5 

```{r}
ggplot(corruptDF,
       mapping = aes(x = CPI2017, y = HDI2017,
                     color = region, 
                     fill = region)) +
  geom_point() +
  geom_smooth(method = "gam") +
  facet_wrap(~ region)
```

I changed used the facet_wrap function to show each region's relationship between HDI2017 and CPI2017.

## Question 6

```{r}
ggplot(corruptDF,
       mapping = aes(x = CPI2017, y = HDI2017,
                     color = region, 
                     fill = region)) +
  geom_point() +
  geom_smooth(method = "gam") +
  scale_x_reverse() +
  facet_wrap(~ region)
```

Using the scale_x_reverse function, I reversed the scale mapping.

## Question 7

```{r}
ggplot(corruptDF,
       mapping = aes(x = CPI2017, y = HDI2017,
                     color = region, 
                     fill = region)) +
  geom_point() +
  geom_smooth(method = "gam") +
  labs(title = "Relationship Btwn Corruption Perceptions and Human Development Index Score in 2017", 
       subtitle = "Data points are countries",
       caption = "Source: Transparency.org/cpi ") +
  scale_x_reverse() +
  facet_wrap(~ region)
```

## Question 8

```{r}
ggsave(filename= "CPIxHDI_figure.pdf")
```
